{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-03-05 thesis/landslide.tex
summarize need for bpor

@@

Iterative Context Bounding \cite{chess-icb} is a state space exploration strategy
that prioritizes interleavings with fewer total preemptions first.
Let $P(I)$ denote the number of preemptions in an interleaving $I$.
Then, to summarize in pseudocode a na\"ive exploration of some state space $S$ as:

\begin{algorithm}[h]
	\ForEach{$I \in S$}{
		Execute($I$)
	}
	\caption{Straightforward exploration ordering}
	\label{alg:not-icb}
\end{algorithm}

@@

\begin{algorithm}[h]
	\For{$n \in [0..\mathsf{max}_P(S)]$}{
		\ForEach{$I \in S, P(I) \le n$}{
			Execute($I$)
		}
	}
	\caption{ICB exploration ordering}
	\label{alg:icb}
\end{algorithm}

%% stupid imperative way to write it
%\begin{algorithm}[h]
%	\For{$n \in [0..\infty]$}{
%		$\mathsf{skipped\_any} := \mathsf{false}$ \\
%		\ForEach{$I \in S$}{
%			\uIf{$P(I) \le n$}{
%				Execute($I$)
%			} \Else {
%				$\mathsf{skipped\_any} := \mathsf{true}$
%			}
%		}
%		\If{$\neg \mathsf{skipped\_any}$}{
%			{\bf Break}
%		}
%	}
%\end{algorithm}

First of all, note that Algorithm~\ref{alg:icb} is structured in a way that repeats interleavings
with fewer than $n$ preemptions that have already been checked in previous iterations of the outer loop.
This is because the number of preemptions in each branch is not known in advance;
rather, the state space must always be explored in an overall depth-first approach,
at best skipping too-preemptful interleavings as they are encountered.
As simple as it would be to state ``{\bf foreach} $I \in {\mathsf{sort}_P(S)}$'' in pseudocode,
implementing such an ordering would be much less straightforward. %
% TODO: post committee review
%\footnote{Translating conference papers' pseudocode into usable implementations is often quite difficult \cite{itg2}.}
The core ICB loop appears in {\tt time\_travel()} in {\tt landslide.c}.
%(although not explicitly structured as a C-style loop in the code)

Secondly, prior work \cite{bpor} has shown that
when combined with DPOR to prune equivalent interleavings,
DPOR's reduction might not be sound with respect to the subset of $S$ under $P \le n$.
That work introduced Bounded Partial Order Reduction (BPOR),
a compatibility extension to DPOR for ICB to address this problem.
To summarize, % TODO: summarize this once you write the DPOR section so you know what terminology to write in terms of
%I refer the reader interested in the algorithm's details to the citation;
%this section will discuss Landslide's implementation.
% TODO: say how u do it

>>>>

% TODO

@@

Iterative Context Bounding \cite{chess-icb} is a state space exploration strategy
that prioritizes interleavings with fewer total preemptions first.
Let $P(S)$ denote the number of preemptions in an execution sequence $S$.
Then, to summarize in pseudocode a na\"ive exploration of some state space $U$ as:

\begin{algorithm}[h]
	\ForEach{$S \in U$}{
		Execute($S$)
	}
	\caption{Straightforward exploration ordering}
	\label{alg:not-icb}
\end{algorithm}

@@

\begin{algorithm}[h]
	\For{$n \in [0..\mathsf{max}_P(U)]$}{
		\ForEach{$S \in U, P(S) \le n$}{
			Execute($S$)
		}
	}
	\caption{ICB exploration ordering}
	\label{alg:icb}
\end{algorithm}

%% stupid imperative way to write it
%\begin{algorithm}[h]
%	\For{$n \in [0..\infty]$}{
%		$\mathsf{skipped\_any} := \mathsf{false}$ \\
%		\ForEach{$S \in U$}{
%			\uIf{$P(S) \le n$}{
%				Execute($S$)
%			} \Else {
%				$\mathsf{skipped\_any} := \mathsf{true}$
%			}
%		}
%		\If{$\neg \mathsf{skipped\_any}$}{
%			{\bf Break}
%		}
%	}
%\end{algorithm}

First of all, note that Algorithm~\ref{alg:icb} is structured in a way that repeats interleavings
with fewer than $n$ preemptions that have already been checked in previous iterations of the outer loop.
This is because the number of preemptions in each branch is not known in advance;
rather, the state space must always be explored in an overall depth-first approach,
at best skipping too-preemptful interleavings as they are encountered.
As simple as it would be to state ``{\bf foreach} $S \in {\mathsf{sort}_P(U)}$'' in pseudocode,
implementing such an ordering would be much less straightforward. %
% TODO: post committee review
%\footnote{Translating conference papers' pseudocode into usable implementations is often quite difficult \cite{itg2}.}
The core ICB loop appears in {\tt time\_travel()} in {\tt landslide.c}.
%(although not explicitly structured as a C-style loop in the code)

\subsubsection{Bounded Partial-Order Reduction}

Secondly, prior work \cite{bpor} has shown that
when combined with DPOR to prune equivalent interleavings,
DPOR's reduction might not be sound with respect to the subset of $U$ under $P \le n$.
That work introduced Bounded Partial Order Reduction (BPOR),
a compatibility extension to DPOR for ICB to address this problem.

To summarize, when DPOR identifies some interleaving $I_{ij}$ to test,
it may not be possible to execute $T(t_j)$ after $t_i$ without exceeding the current preemption bound.
However, there may exist another interleaving $J_{i'j}$ within the bound which runs $T(t_j)$ before $t_i$.
If a DPOR implementation na\"ively configured with ICB simply skipped $I$ on account of the preemption bound,
there's no guarantee $J$ would get marked for exploration
from any other iteration and/or any other pair of conflicting transitions.
Even though restricting the state space to a certain maximum preemption bound is already unsound
in terms of losing full concurrency coverage,
failing to test even $J$ would be a failure of DPOR itself
to soundly prune the already-reduced state space defined by the bound.
Hence the need for BPOR, to ensure that if such an alternative $J$ to $I$ exists within the bound,
$J$ gets marked for exploration immediately.

%I refer the reader interested in the algorithm's details to the citation;
%this section will discuss Landslide's implementation.
% TODO: say how u do it
